---
output: github_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "png",
  warning = FALSE
)
```

# Wisconet v1 API wrapper for R

`rwisconet` provides an interface to the Wisconet v1 API (<https://wisconet.wisc.edu/api/v1>). Supports fetching station metadata, available field definitions, and observation data for one or more stations.

## Installation

```{r, eval=FALSE}
remotes::install_github("bzbradford/rwisconet")
```

## Basic Usage

### Initialize the client

Creating a new client automatically fetches station and field metadata from the API. Default timezone is "America/Chicago" for interpreting date/time ranges and the conversion of UTC to local time.

```{r}
suppressPackageStartupMessages(library(tidyverse))
library(rwisconet)

# initialize the connection
wn <- Wisconet$new()
```

You can skip the initial metadata fetch or use a different timezone, but normally you would leave the default settings.

```{r, eval=FALSE}
wn <- Wisconet$new(timezone = "US/Eastern", fetch_on_init = FALSE)
```

### Configuration

Several configutation options are available in the `wn$config` object. For example, the timezone is stored in  `wn$config$timezone`. There are several other configuration options you can set that affect the throttling, concurrency, or retry behavior of API calls to Wisconet. Each can be set like `wn$config${config_name} <- {config_value}`, but default values should work fine. See all config options and values by executing `wn` or `wn$print()`:

```{r}
wn
```

### Explore stations

The station metadata table is stored in `wn$stations`. You can refresh it at any time with `wn$get_stations()`.

```{r}
# extract the stations table
stns <- wn$stations

# view it
glimpse(stns)
```

Find the nearest station(s) to a given point:

```{r}
# will return the nearest 3 stations to this point and the distance
wn$find_nearest_station(lat = 43.07, lng = -89.40, n = 3)
```

Display all stations on an interactive map (requires the `leaflet` package):

```{r, eval=FALSE}
library(leaflet)

# create a leaflet map
leaflet(wn$stations) |>
  addTiles() |>
  addMarkers(
    lng = ~longitude,
    lat = ~latitude,
    # on mouse hover points will show station id and name
    label = ~ paste0(station_id, ": ", station_name)
  )
```

### Explore available fields

The fields table is stored in `wn$fields`. Use `find_fields()` to filter by measure type and/or collection frequency:
  
```{r}
# List all available fields
wn$fields

# Calling find_fields with no args shows types and frequency options
wn$find_fields()

# Show air temp fields
wn$find_fields(type = "Air Temp")

# show daily fields
wn$find_fields(freq = "daily")

# show air temp fields with 5 minute collection frequency
wn$find_fields(type = "Air Temp", freq = "5min")
```

### Fetch observations

#### Single station

Use `get_measures()` to fetch data for one station. Provide a station ID, a vector of field `standard_name` values, and a time window. Returned data is formatted as `tibble` with station and measure metadata. 

```{r}
# fields to download
measures_5min <- c(
  "5min_air_temp_f_avg",
  "5min_relative_humidity_pct_avg",
  "5min_wind_speed_mph_avg"
)

# run the query
obs <- wn$get_measures(
  stn_id = "HNCK", # Hancock
  fields = measures_5min,
  start_time = now() - days(1),
  end_time = now()
)

# view results
glimpse(obs)
```

#### Multiple stations

Use `get_measures_stations()` to fetch data for a specific set of stations in parallel:

```{r}
# run the query
obs <- wn$get_measures_stations(
  stn_ids = c("ALTN", "HNCK"), # Arlington, Hancock
  fields = measures_5min,
  start_time = now() - days(1),
  end_time = now()
)

# view results
glimpse(obs)
```

#### Specific start/end times per station

You can also pass a named list of per-station start and end times if you need to collect data for different timeframes by station.

```{r}
# fields to download
measures_daily <- c(
  "daily_air_temp_f_min",
  "daily_air_temp_f_avg",
  "daily_air_temp_f_max"
)

# set start times
starts <- list(
  ALTN = now() - months(2),
  HNCK = now() - months(1)
)

# set end times for each station 7 days from start time
ends <- lapply(starts, \(x) x + days(7))

# run the query
obs <- wn$get_measures_stations(
  stn_ids = names(starts),
  fields = measures_daily,
  start_time = starts,
  end_time = ends
)

# show results
glimpse(obs)
```

#### All active stations

Use `get_measures_all()` to query every active station. Note that this will hit the API throttle and may take a minute to resolve.

```{r}
# run the query
obs <- wn$get_measures_all(
  fields = measures_daily,
  start_time = today() - days(3),
  end_time = today()
)

# show results
glimpse(obs)
```

### Handling received data

You can simplify the returned data by selecting only a few important data columns.

```{r}
# run the query
obs <- wn$get_measures_stations(
  stn_ids = c("ALTN", "HNCK"), # Arlington, Hancock
  fields = measures_daily,
  start_time = now() - months(1),
  end_time = now()
)

# select a few specific columns
select_obs <- obs |>
  select(station_id, dttm_local, date, standard_name, measure_value)

# pivot to wide format
obs_wide <- select_obs |>
  pivot_wider(names_from = standard_name, values_from = measure_value)

# view data
obs_wide
```

Generate a simple figure from the returned data:

```{r fig}
# plot the data (in long form) from above
select_obs |>
  ggplot(aes(x = dttm_local, y = measure_value, color = standard_name)) +
  geom_line() +
  facet_wrap(~station_id, ncol = 1) +
  theme(legend.position = "bottom")
```

## Development

Common tasks:

```{r dev, eval=FALSE}
# check package
devtools::check()

# build package
devtools::build()

# run tests
testthat::test_dir("tests")

# build documentation
devtools::document()

# build readme from Rmd
devtools::build_readme()

# build a pdf manual
devtools::build_manual()
```
